---
title: "Olympic Trial Running"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Checks on Data
Data is running olympic trial data - 200,400,800, and 1500m runs only. 

First, run Leven's Test of Homogeneity and D'Agostino Skewness Test on the model (Percent Difference by Stroke, Year, and Distance).

```{r, echo = FALSE,message = FALSE}
library(dplyr); library(moments); library(car);
# Read in data, remove all NA, convert % diff to numeric, remove any 0's, and convert year to factor
ot.data <- read.csv("~/OneDrive/Research/Swimming/runningAnalysis/otRunningData.csv") %>%
subset(distance == 100 | distance == 200 | distance == 400 | distance == 800 | distance == 1500) %>% subset(year > 1970) %>% mutate(interaction = interaction(as.factor(place), distance)) %>% subset(interaction != "8.100")
ot.data[is.na(ot.data)] <- 0
ot.data$per.diff <- as.numeric(as.character(ot.data$per.diff))
ot.data <- subset(ot.data, per.diff > .1)
ot.data$year <- as.factor(ot.data$year)

# Using a 2 factor anova with fixed effects: Year & Distance
leveneTest(per.diff ~ year*distance, data = ot.data, center = mean)
agostino.test(ot.data$per.diff)
```

This, therefore requires non parametric testing. For this, Kruskal-Wallis Rank Sum Test with Pairwise Wilcoxon post hocs will be used.

### Main effect of year
The KW result will be first followed by the Wilcoxon post hoc tests with Holm-Bonferroni corrections (value = p value for given comparison).

```{r, echo = F, warning = F, message = F}
library(reshape)
ot.data %>% group_by(year) %>% summarize(mean = mean(per.diff, na.rm = TRUE), sd = sd(per.diff, na.rm = TRUE), median(per.diff))
kruskal.test(per.diff ~ year, data = ot.data)
  
yr.ph <- pairwise.wilcox.test(ot.data$per.diff, ot.data$year, paired = F, p.adjust.method="holm") #post hocs
lapply(as.list(seq(1968, 2016, 4)), function(x) {subset(melt(yr.ph[[3]]), X1 == x | X2 == x) %>% subset(value > 0) %>% round(6)})
```

### Main Effect for Distance 
The KW result will be first followed by the Wilcoxon post hoc tests with Holm-Bonferroni corrections (value = p value for given comparison).
```{r, echo = F, warning = F}
ot.data %>% group_by(distance) %>% summarize(mean = mean(per.diff, na.rm = TRUE), sd = sd(per.diff, na.rm = TRUE), median(per.diff))
kruskal.test(per.diff ~ distance, data = ot.data)
  
d.ph <- pairwise.wilcox.test(ot.data$per.diff, ot.data$distance, paired = F, p.adjust.method="holm") #post hocs
lapply(as.list(c(100, 200, 400, 800, 1500)), function(x) {subset(melt(d.ph[[3]]), X1 == x | X2 == x) %>% subset(value > 0) %>% round(6)})
```

\newpage # Figures

Figure 1: Gender gap for each year and event

```{r, echo = FALSE, warning  = FALSE}
library(ggplot2)
data2 <- ot.data
data2$year <- as.factor(data2$year)
data2$distance <- as.factor(data2$distance)
 
year.means <- data2 %>% 
  group_by(year) %>% 
  summarize(
        mean = mean(per.diff, na.rm = TRUE), 
        sd = sd(per.diff, na.rm = TRUE),
        max = max(per.diff), 
        min = min(per.diff))

########## Year Plot ##########

# Manipulate data for graphing
data2$year <- as.numeric(data2$year) 

# Make new Year means table with similar x axis (Percent_Difference)
names(year.means) <- c('Year','Percent_Difference','SD','Max','Min')
year.means$Year <- as.numeric(year.means$Year) 

# Theme for Plot
theme <- theme(
               plot.background = element_rect(fill = "transparent",color = NA),
               panel.background = element_rect(fill = "transparent",color = NA),
               axis.line.x = element_line(size = .5, color = "black"),
               axis.line.y = element_line(size = .5, color = "black"),
               axis.text = element_text(family = "Times",size = 16, color = "black"),
               axis.title.x = element_blank(),
               axis.title.y = element_text(family = "Times", size = 16, margin = margin(r = 10)),
               axis.ticks.length = unit(.25, "cm"),
               axis.ticks = element_line(colour = 'black', size = .5),
               legend.text = element_text(size = 14, family = "Times"),
               legend.title = element_blank(),
               legend.key = element_rect(fill = "white"),
               plot.margin=unit(c(1,1,1,1),"cm")
)

# Define shapes and breaks
distance.shapes = c(21,24,23,22,21)

# Begin Construct Graph
base <- ggplot(data = data2, aes(x = year, y = per.diff))
scatter <- geom_point(size = 5, position = position_jitter(width = 0.8), aes(shape = distance, fill = distance), color = "black", stroke = 1) 
line <- geom_line(data = year.means, aes(x = Year, y = Percent_Difference), size = 3)
ribbon <- geom_ribbon(data = year.means, aes(x = Year, y = Percent_Difference, ymin = Percent_Difference-SD,ymax = Percent_Difference+SD),alpha = 0.15) 

# Combine for graph
raw <- base + scatter + theme +
scale_shape_manual(values = distance.shapes, 
                   guide = guide_legend(keywidth=.2,keyheight=.28,default.unit="inch"),
                   breaks=c("100","200","400", "800","1500")) + 
  scale_fill_manual(values = c("white","#888888","#222222","#444444", "#666666"),
                    breaks=c("100","200","400", "800","1500")) +
  line + 
  ribbon + 
  scale_x_discrete(limits = c("1972","1976","1980","1984","1988","1992",
                              "1996","2000","2004","2008","2012","2016")) +
  ylab("Performance Gap (% Difference)") +
  scale_y_continuous(breaks = c(seq(6,24,2))) + 
  expand_limits(y = c(6,24), x = c(.5,12.5)) + 
  annotate("text", label = "A", size = 12, x = 1, y = 24.5, family = "Times")

plot(raw)
```

\newpage Figure 2: LOESS Smoothing Plot for each year and distance

```{r, echo = F, warning= FALSE}
smooth <- ggplot(data = ot.data, aes(x = as.numeric(year), y = per.diff)) + geom_smooth(size = 1.2, color = "black") + 
  geom_smooth(size = .5, color = "black", alpha = 0.15, aes(linetype = distance)) +
  theme +
  ylab("Performance Gap (% Difference)") +
  scale_x_discrete(limits = c("1972","1976","1980","1984","1988","1992",
                              "1996","2000","2004","2008","2012","2016")) +
  scale_y_continuous(breaks = c(seq(8,24,4))) + expand_limits(y = c(8,24), x = c(.5,12.5)) + 
  scale_linetype_manual(values = c(3,2,4,5,6), guide = guide_legend(keywidth = .5, keyheight=.3,default.unit="inch"), breaks=c("100","200","400", "800","1500")) +
  annotate("text", label = "B", size = 12, x = 1, y = 24.5, family = "Times")

plot(smooth)
```